---
title: "Final Project"
output: html_document
---

An attempt to find the optimal number of calls to be made on an account in a month by a sales representative.

1) Import data
2) Scrub data
3) Prepare data
4) Explore data
5) Analyze data
6) Draw conclusions

###Librairies used
```{r}
#install.packages("date")
#install.packages("zipcode")
#install.packages("ggplot2"")
#install.packages("ggmap")

library(dplyr) #select arrange
library(date)  #as.date
library(zipcode) #adds leading zipcodes
library(ggplot2)
library(ggmap)
library(RCurl)
#library(plyr)   #rename columns
```

### Import Data stored on GitHub 
```{r}
mydata <- NULL
data <- getURL("https://raw.githubusercontent.com/srkhan123/Final-Project/master/skd.csv",
               ssl.verifypeer=0L, followlocation=1L)
mydata <- read.csv(text=data)
str(mydata)
head(mydata)
```

### Tidy up Data
```{r}
mydata$TG_ID <- factor(mydata$TG_ID) #TG_ID should by factor as an identifier
mydata$ZIPCODE <- as.factor(clean.zipcodes(mydata$ZIP)) #clean zipcodes by adding leading 0 and 5 digit long
mydata$ZIP <- NULL #delete old Zip column
head(mydata)
str(mydata)

```

### Add map coordinates to zipcodes Longitude and Latitude for ggmap
```{r}
data(zipcode)
mydata <- merge(mydata, zipcode, by.x = 'ZIPCODE', by.y = 'zip', all.x=TRUE)
head(mydata)

```


#Histogram of Units 
```{r}

myunits <- mydata %>%
          filter(UNITS != 0)

#myunits[ which(myunits$UNITS == 0),]  

hist(myunits$UNITS, breaks = 50, main="Order Size Frequency", xlab="UNITS", col="lightgreen", xlim = c(-20,100))

```


#Histogram of CAlls 
```{r}

mycalls <- mydata %>%
          filter(CALLS != 0)

#mycalls[ which(mycalls$CALLS == 0),]  

hist(mycalls$CALLS, breaks = 50, main="Number of Call Frequency", xlab="CALLS", col="lightblue", xlim = c(0,10))

```

#Plot and exploring Correlation
```{r}

plot(mydata$CALLS, mydata$UNITS , xlab="CALLS" , ylab="UNITS")

cor(myunits$CALLS, myunits$UNITS)

cor(mycalls$CALLS, mycalls$UNITS)

```

#Quantile Plot
```{r}
qqnorm(mydata$UNITS, main = "Normal Q-Q Plot for Units")
qqline(mydata$UNITS)

qqnorm(mydata$CALLS, main = "Normal Q-Q Plot for Calls")
qqline(mydata$CALLS)

```

#Plotting data for zipcodes from the package using ggplot2
```{r}
data(zipcode)
zipcode$region = substr(zipcode$zip, 1, 1)
 
g = ggplot(data = zipcode) + geom_point(aes(x=longitude, y=latitude, colour=region))
 
# simplify display and limit to the "lower 48"
g = g + theme_bw() + scale_x_continuous(limits = c(-125,-66), breaks = NULL)
g = g + scale_y_continuous(limits = c(25,50), breaks = NULL)
 
# don't need axis labels
g = g + labs(x=NULL, y=NULL)
g

```

#Plotting using mydata and ggplot2
```{r}
g <- NULL
g = ggplot(data = mydata) 
g = g + geom_point(aes(x=CALLS, y=UNITS))
#g = g + facet_grid(. ~ REGION)
#g = g + geom_smooth(method = "lm")
g = g + theme_gray()
g

```


#using ggmap with Units
```{r}
map <- NULL
map <- get_map(location = 'united states', zoom = 4, maptype = "terrain", 
               source = 'google', color = 'color')

ggmap(map)+geom_point(aes (x=longitude, y=latitude, show_guide = TRUE,color = UNITS), 
                      data = myunits, alpha = 0.5, na.rm = T )
     
```


#using ggmap with Calls
```{r}
map <- NULL
map <- get_map(location = 'united states', zoom =4, maptype = "terrain", 
               source = 'google', color = 'color')

ggmap(map)+geom_point(aes (x=longitude, y=latitude, show_guide = TRUE,color = CALLS), 
                      data = mycalls, alpha = 0.5, na.rm = T )
      
```



#Normalizing data to show units for each account over 3 years
```{r}
str(mydata)
myaccounts <- NULL
myaccounts <- mydata %>%
              group_by(ACCOUNT_NUM, NAME, ADDRESS, CITY, STATE, ZIPCODE) %>% #
              #select(OUTLET_NUM, NAME, ADDRESS, CITY, STATE, ZIPCODE) %>%
              summarise( 
                TOTAL_UNITS = sum(UNITS), 
                TOTAL_CALLS = sum(CALLS)
              ) 


#check an account
myaccounts[ which(myaccounts$TOTAL_UNITS >0),]   

myaccounts[ which(myaccounts$ACCOUNT_NUM == '15143201'),]   

mydata[ which(mydata$ACCOUNT_NUM == '15143201'),]   
                   
       

head (myaccounts)
str(myaccounts)
              
summary(myaccounts)
  

```


#adding number of counts for units > 0 and Calls > 0
```{r}

UnitsCount <- NULL
UnitsCount <- mydata %>% 
              filter(UNITS > 0) %>%
              count(ACCOUNT_NUM) 
colnames(UnitsCount)[colnames(UnitsCount)=="n"] <- "NumberOfMonthsOrdered"


myaccountUnits <- merge(myaccounts, UnitsCount, by.x = 'ACCOUNT_NUM', by.y = 'ACCOUNT_NUM' , all.x=TRUE)

        
#check an account
myaccountUnits[ which(myaccountUnits$ACCOUNT_NUM == '15143201'),]   

#----------------------------------------------------------------------------------------------------

CallsCount <- NULL
CallsCount <- mydata %>% 
              filter(CALLS > 0) %>%
              count(ACCOUNT_NUM) 

colnames(CallsCount)[colnames(CallsCount)=="n"] <- "NumberOfMonthsCalled"

 
#check an account
CallsCount[ which(CallsCount$ACCOUNT_NUM == '15143201'),]

myaccounts <- merge(myaccounts, CallsCount, by.x = 'ACCOUNT_NUM', by.y = 'ACCOUNT_NUM' , all.x=TRUE)
          
#check an account
myaccounts[ which(myaccounts$ACCOUNT_NUM == '15143201'),]

#-------------------

myaccounts <- merge(myaccounts, myaccountUnits, by.x = 'ACCOUNT_NUM', by.y = 'ACCOUNT_NUM' , all.x=TRUE)


myaccounts[ which(myaccounts$ACCOUNT_NUM == '15143201'),]

myaccounts$NAME.y <- NULL  
myaccounts$ADDRESS.y  <- NULL
myaccounts$CITY.y  <- NULL
myaccounts$STATE.y  <- NULL
myaccounts$ZIPCODE.y  <- NULL
myaccounts$TOTAL_UNITS.y  <- NULL
myaccounts$TOTAL_CALLS.y <- NULL


myaccounts[ which(myaccounts$ACCOUNT_NUM == '15143201'),]


```

#Analyzing data further
```{r}
#graph units and calls

plot(myaccounts$TOTAL_UNITS.x~myaccounts$TOTAL_CALLS.x, xlab = "CALLS", ylab = "UNITS")

myaccounts[ which(myaccounts$TOTAL_UNITS.x > 2000),]

myaccounts[ which(myaccounts$NumberOfMonthsOrdered > 25),]

plot(myaccounts$NumberOfMonthsOrdered~myaccounts$NumberOfMonthsCalled, xlab = "NumberOfMonthsCalled in last 3 years", ylab = "NumberOfMonthsOrdered in last 3 years")

cor(myaccounts$NumberOfMonthsOrdered, myaccounts$NumberOfMonthsCalled)
```

### The aboveplot shows that number of optimum calls in 3 years is 20 to get the most optimum level of ordering
### which is equal to which gives orders about 50% of the time



####Analyzing data further with Quartiles
```{r}

levels = myaccounts$TOTAL_UNITS.x     
quantile(levels, c(0, .25, .5, .75, 1), na.rm = TRUE) 

# 0%  25%  50%  75% 100% 
# -12    6   16   42 2568 

quantile(myaccounts$TOTAL_UNITS.x , probs = seq(0, 1, 0.25), na.rm = TRUE, names = TRUE)

myQuantiles <- NULL
myQuantiles <- within(myaccounts, quantile <- cut(myaccounts$TOTAL_UNITS.x, quantile(myaccounts$TOTAL_UNITS.x, probs=0.25, 0.5, 0.75), include.lowest=TRUE, labels=FALSE))

myQuantiles[ which(myQuantiles$NumberOfMonthsOrdered > 25),]

plot(myQuantiles$quantile~myQuantiles$NumberOfMonthsOrdered, xlab = "NumberOfMonthsOrdered", ylab="Quantiles")


plot(myQuantiles$quantile~myQuantiles$NumberOfMonthsCalled, xlab = "NumberOfMonthsCalled", ylab = "Quantiles")

```


#Quantile Plot
```{r}
qqnorm(myQuantiles$NumberOfMonthsOrdered, main = "Normal Q-Q Plot for Units")
qqline(myQuantiles$NumberOfMonthsOrdered)

qqnorm(myQuantiles$NumberOfMonthsCalled, main = "Normal Q-Q Plot for Calls")
qqline(myQuantiles$NumberOfMonthsCalled)

```



#Linear Regression
```{r}
lmResult <- with(myaccounts, lm(NumberOfMonthsOrdered ~ NumberOfMonthsCalled))

summary(lmResult)


plot(myaccounts$NumberOfMonthsOrdered~myaccounts$NumberOfMonthsCalled, xlab = "NumberOfMonthsCalled in last 3 years", ylab = "NumberOfMonthsOrdered in last 3 years")
abline(lmResult, col="red")

plot(lmResult)

```

###Conclusion
##### the Plot data shows with abline that you need to make about 2 calls to get an order on an account